---
title: "Corey"
author: "https://github.com/jrkkv5/FIN6572Fall2020/blob/master/National%20Debt/importNationalDebt.Rmd"
date: "11/8/2020"
output: html_document
---

``` {r basic analysis}
library(readr)
library(devtools)
library(blsAPI)
library(rjson)
library(curl)
library(dplyr)
library(tidyverse)
#Loads necessary packages


merged_final <- read_csv("csv files/mergedFinal.csv")

#Update code for your states and run regression from 1984-2014, create some basic plots, etc.

```
`````` {r MN Basic Analysis}

MN_base <- merged_final[ merged_final$state == "MN", ]

scatter.smooth(MN_base$year, MN_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue")

scatter.smooth(MN_base$year, MN_base$HPI, main="HPI by Year", xlab = "Year", ylab = "HPI", col = "blue")

scatter.smooth(x=MN_base$year, y=MN_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "red")

scatter.smooth(x=MN_base$year, y=MN_base$poverty, main="% of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20))

MN_base$RMHI <- as.integer(MN_base$RMHI)
plot(x=MN_base$year, y=MN_base$RMHI, main="Median Household Income by Year", xlab = "Year", ylab = "RMHI" )

MN_base$population <- as.integer(MN_base$population)
plot(x=MN_base$year, y=MN_base$population, main="Population by Year", xlab = "Year", ylab = "Population" )
 
#test to show a scatterplot MN HPI data against year
```
`````` {r AL Basic Analysis}

AL_base <- merged_final[ merged_final$state == "AL", ]

scatter.smooth(AL_base$year, AL_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue")

scatter.smooth(AL_base$year, AL_base$HPI, main="HPI by Year", xlab = "Year", ylab = "HPI", col = "blue")

scatter.smooth(x=AL_base$year, y=AL_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "red")

scatter.smooth(x=AL_base$year, y=AL_base$poverty, main="% of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20))

AL_base$RMHI <- as.integer(AL_base$RMHI)
plot(x=AL_base$year, y=AL_base$RMHI, main="Median Household Income by Year", xlab = "Year", ylab = "RMHI" )

AL_base$population <- as.integer(AL_base$population)
plot(x=ALbase$year, y=AL_base$population, main="Population by Year", xlab = "Year", ylab = "Population" )
 
#test to show a scatterplot MN HPI data against year
```
`````` {r LA Basic Analysis}

LA_base <- merged_final[ merged_final$state == "LA", ]

scatter.smooth(LA_base$year, LA_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue")

scatter.smooth(LA_base$year, LA_base$HPI, main="HPI by Year", xlab = "Year", ylab = "HPI", col = "blue")

scatter.smooth(x=LA_base$year, y=LA_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "red")

scatter.smooth(x=LA_base$year, y=LA_base$poverty, main="% of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20))

LA_base$RMHI <- as.integer(LA_base$RMHI)
plot(x=LA_base$year, y=LA_base$RMHI, main="Median Household Income by Year", xlab = "Year", ylab = "RMHI" )

LA_base$population <- as.integer(LA_base$population)
plot(x=LA_base$year, y=LA_base$population, main="Population by Year", xlab = "Year", ylab = "Population" )
 
#test to show a scatterplot MN HPI data against year
```