---
title: "Corey"
author: "https://github.com/jrkkv5/FIN6572Fall2020/blob/master/National%20Debt/importNationalDebt.Rmd"
date: "11/8/2020"
output: html_document
---

``` {r basic analysis}
library(readr)
library(devtools)
library(blsAPI)
library(rjson)
library(curl)
library(dplyr)
library(tidyverse)
#Loads necessary packages

merged_final <- read_csv("csv files/mergedFinal.csv")
#Reads final file into merged dataframe

merged_final$log_pop = log(merged_final$population)

merged_final$log_RMHI = log(merged_final$RMHI)
#Update code for your states and run regression from 1984-2014, create some basic plots, etc.

```

`````` {r CO Basic Analysis}

CO_observations <- merged_final[ merged_final$state == "CO", ]
CO_base <- CO_observations[ CO_observations$year < 2014, ] # Creates training set for model
CO_sampleset <-  CO_observations[ CO_observations$year > 2013, ] # Creates set to test model

scatter.smooth(CO_base$year, CO_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue")

scatter.smooth(CO_base$year, CO_base$HPI, main="HPI by Year", xlab = "Year", ylab = "HPI", col = "blue")

scatter.smooth(x=CO_base$year, y=CO_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "red")

scatter.smooth(x=CO_base$year, y=CO_base$poverty, main="% of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20))

CO_base$RMHI <- as.integer(CO_base$RMHI)
plot(x=CO_base$year, y=CO_base$RMHI, main="Median Household Income by Year", xlab = "Year", ylab = "RMHI" )

CO_base$population <- as.integer(CO_base$population)
plot(x=CO_base$year, y=CO_base$population, main="Population by Year", xlab = "Year", ylab = "Population" )


cor(CO_base$RMHI, CO_base$value)
cor(CO_base$HPI, CO_base$value)
cor(CO_base$poverty, CO_base$value)
cor(CO_base$population, CO_base$value)
cor(CO_base$sp500, CO_base$value)
#run base correlations between categories and dependent variable

par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(CO_base$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(CO_base$value)$out))  # box plot for 'Unemployment'
boxplot(CO_base$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(CO_base$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level


CO_Test <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = CO_base)
summary(CO_Test)
anova(CO_Test)
#run multiple linear model for data

CO_linearModelSignificant <- lm(value ~ poverty + population, data = CO_base)
summary(CA_linearModelSignificant)
#test to show a scatterplot CA HPI data against year

CO_linearModel <- lm(value ~ poverty + RMHI + HPI + log(population) + sp500, data = CO_base)
summary(CO_linearModel)
anova(CO_linearModel)
```
`````` {r OR Basic Analysis}
OR_observations <- merged_final[ merged_final$state == "OR", ]
OR_base <- OR_observations[ OR_observations$year < 2014, ] # Creates training set for model
OR_sampleset <-  OR_observations[ OR_observations$year > 2013, ] # Creates set to test model

scatter.smooth(OR_base$year, OR_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue")

scatter.smooth(OR_base$year, OR_base$HPI, main="HPI by Year", xlab = "Year", ylab = "HPI", col = "blue")

scatter.smooth(x=OR_base$year, y=OR_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "red")

scatter.smooth(x=OR_base$year, y=OR_base$poverty, main="% of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20))

OR_base$RMHI <- as.integer(OR_base$RMHI)
plot(x=OR_base$year, y=OR_base$RMHI, main="Median Household Income by Year", xlab = "Year", ylab = "RMHI" )

OR_base$population <- as.integer(OR_base$population)
plot(x=OR_base$year, y=OR_base$population, main="Population by Year", xlab = "Year", ylab = "Population" )


cor(OR_base$RMHI, OR_base$value)
cor(OR_base$HPI, OR_base$value)
cor(OR_base$poverty, OR_base$value)
cor(OR_base$population, OR_base$value)
cor(OR_base$sp500, OR_base$value)
#run base correlations between categories and dependent variable

par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(OR_base$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(OR_base$value)$out))  # box plot for 'Unemployment'
boxplot(OR_base$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(OR_base$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level


OR_Test <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = OR_base)
summary(OR_Test)
anova(OR_Test)
#run multiple linear model for data

OR_linearModelSignificant <- lm(value ~ poverty + population, data = OR_base)
summary(OR_linearModelSignificant)
#test to show a scatterplot OR HPI data against year

OR_linearModel <- lm(value ~ poverty + RMHI + HPI + log(population) + sp500, data = OR_base)
summary(OR_linearModel)
anova(OR_linearModel)
```

`````` {r CT Basic Analysis}

CT_observations <- merged_final[ merged_final$state == "CT", ]
CT_base <- CT_observations[ CT_observations$year < 2014, ] # Creates training set for model
CT_sampleset <-  CO_observations[ CT_observations$year > 2013, ] # Creates set to test model

scatter.smooth(CT_base$year, CT_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue")

scatter.smooth(CT_base$year, CT_base$HPI, main="HPI by Year", xlab = "Year", ylab = "HPI", col = "blue")

scatter.smooth(x=CT_base$year, y=CT_base$value, main="Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "red")

scatter.smooth(x=CT_base$year, y=CT_base$poverty, main="% of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20))

CT_base$RMHI <- as.integer(CT_base$RMHI)
plot(x=CT_base$year, y=CT_base$RMHI, main="Median Household Income by Year", xlab = "Year", ylab = "RMHI" )

CT_base$population <- as.integer(CT_base$population)
plot(x=CT_base$year, y=CT_base$population, main="Population by Year", xlab = "Year", ylab = "Population" )


cor(CT_base$RMHI, CT_base$value)
cor(CT_base$HPI, CT_base$value)
cor(CT_base$poverty, CT_base$value)
cor(CT_base$population, CT_base$value)
cor(CT_base$sp500, CT_base$value)
#run base correlations between categories and dependent variable

par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(CT_base$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(CT_base$value)$out))  # box plot for 'Unemployment'
boxplot(CT_base$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(CT_base$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level


CT_Test <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = CT_base)
summary(CT_Test)
anova(CT_Test)
#run multiple linear model for data

CT_linearModelSignificant <- lm(value ~ poverty + population, data = CT_base)
summary(CA_linearModelSignificant)
#test to show a scatterplot CT HPI data against year

CT_linearModel <- lm(value ~ poverty + RMHI + HPI + log(population) + sp500, data = CT_base)
summary(CT_linearModel)
anova(CT_linearModel)
```
















IL Graphs and Histograms 

```{r IL Histogram Unemployment}
hist(IL_data$value,
     main="Histogram of Unemployment Across All Years
     ",
     xlim=c(4,11),
     xlab="Unemployment Rate",
     breaks = 5)
```
Histograms graphically summarize the distribution of a data set. In Illinois, from 1984 to 2018, The annual unemployment rates ranged from 4.0-11.0%. The histogram displays a positively skewed distribution. An unemployment rate that fell within the first two intervals, 4.0-4.9% & 5.0-5.9%, occurred 18 out of the 36 years examined. 

```{r IL Histogram Poverty}
hist(IL_data$poverty,
     main="Histogram of Poverty Across All Years",
     xlim=c(9,16),
     xlab="Poverty Rate",
     breaks = 5)
```
In Illinois, from 1984 to 2018, the poverty levels ranged from 9.0-16.0%. The histogram displays a relatively normal distribution. Between 12.0-12.9% was the most frequently recorded interval during this time period. In 14 of the 36 years examined, the poverty rate in Illinois was between 13-15.9%. 

```{r IL scatter unemployment}
scatter.smooth(IL_data$year, IL_data$value, main="IL Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue", span = 2/8)
```
A scatterplot displays the relationship between two numeric variables. This plot demonstrates the relationship between the annual unemployment rate and the year, from 1984 to 2018. In Illinois, there is no linear relationship, however, there is an observable cyclical pattern. The annual unemployment rate was lowest around 1999 and peaked in 2008. 
```{r IL scatter HPI}
scatter.smooth(IL_data$year, IL_data$HPI, main="IL HPI by Year", xlab = "Year", ylab = "HPI", col = "blue", span = 2/8)
```
This plot represents the relationship between Illinois' median housing price index (HPI) and the year. From 1984 to 2008 the relationship appeared to be linear until the HPI dipped in 2008-2013. Since then, the HPI began to continue its increase in a linear fashion. This pattern is observed consistently when examining many other states and, through inference, can partially be attributed to the housing crisis that occurred in 2008 and 2009. 
```{r IL scatter poverty}
scatter.smooth(x=IL_data$year, y=IL_data$poverty, main="IL % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```
The percentage of individuals living below the poverty rate in Illinois does not display a clear pattern. Overall, there has been slight decrease of roughly 4% from 1984 to 2018. Poverty levels peaked in 1992.
```{r IL scatter log RMHI}
#IL_data$log_RMHI <- as.integer(IL_data$log_RMHI)
scatter.smooth(x=IL_data$year, y=IL_data$log_RMHI, main="IL Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```
The median household income in Illinois increased dramatically from 1984-2018. The lowest and highest values recorded fell at the beginning and the end of the time period, respectively.
```{r IL scatter log population}
#IL_data$log_pop <- as.integer(IL_data$log_pop)
scatter.smooth(x=IL_data$year, y=IL_data$log_pop, main="IL Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```
The log population of Illinois decreased from 1984-1988 before increasing in a relatively linear fashion from 1988 to its peak in 2013. 2013-2018 is characterized by a gradual decline in population. 
```{r IL boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(IL_data$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(IL_data$value)$out))  # box plot for 'Unemployment'
```
The annual unemployment rate in Illinois ranged from 4%-11% from 1984-2018. The boxplot displays a positively skewed distribution meaning the mean is greater than the median. 
```{r IL boxplot poverty}
boxplot(IL_data$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(IL_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```
The poverty levels in Illinois ranged from roughly 9% to just under 15% from 1984-2018. The median poverty rate is displayed at roughly 12.5%.The plot displays a relatively normal distribution meaning the mean and median are close the same value. 




TX Graphs and Histograms
```{r TX Analysis}
TX_data <- merged_final[ merged_final$state == "TX", ]
```

```{r TX Histogram Unemployment}
hist(TX_data$value,
     main="Histogram of Unemployment Across All Years",
     xlim=c(3,9),
     xlab="Unemployment Rate",
     breaks = 5)
```
In Texas, from 1984 to 2018, the annual unemployment rates ranged from 3.0-8.9%. The most common interval, or range of annual unemployment rates was between 5.0-5.9% which occurred 10 times in the 36 year period. Texas only experienced one year with an unemployment rate of 8.0% or higher. 
```{r TX Histogram Poverty}
hist(TX_data$poverty,
     main="Histogram of Poverty Across All Years",
     xlim=c(13,20),
     xlab="Poverty Rate",
     breaks = 5)
```
In Texas, from 1984 to 2018, the annual poverty rates ranged from 13.0-20.0%. The histogram is nearly symmetrical meaning most of the values occurred in the near middle of the range. 27 of the 36 years were characterized by poverty rates ranging from 15.0-17.9%. Only 5 years experienced lower rates and only 4 experienced higher rates of poverty.    
```{r TX scatter unemployment}
scatter.smooth(TX_data$year, TX_data$value, main="TX Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue", span = 2/8)
```
This plot displays the annual unemployment rate in Texas from 1984-2018. There is no linearity to the plot. However, there is a cyclical nature to the unemployment rate over time. The lowest value occurred in 2000. Texas and West Virginia were the only states examined that had a higher unemployment rate before 2008-2009. This occurred in 1986 while the unemployment rate reached over 8%.
```{r TX scatter HPI}
scatter.smooth(TX_data$year, TX_data$HPI, main="TX HPI by Year", xlab = "Year", ylab = "HPI", col = "blue", span = 2/8)
```
The median housing price index displayed a somewhat linear increase from 1984-2008 before flatting out for the next 5 years. By 2013, the positive linear relationship re-established itself. 
```{r TX scatter poverty}
scatter.smooth(x=TX_data$year, y=TX_data$poverty, main="TX % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r TX scatter log RMHI}
#TX_data$log_RMHI <- as.integer(TX_data$log_RMHI)
scatter.smooth(x=TX_data$year, y=TX_data$log_RMHI, main="TX Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r TX scatter log population}
#TX_data$log_pop <- as.integer(TX_data$log_pop)
scatter.smooth(x=TX_data$year, y=TX_data$log_pop, main="TX Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r TX boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(TX_data$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(TX_data$value)$out))  # box plot for 'Unemployment'
```

```{r TX boxplot poverty}
boxplot(TX_data$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(TX_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```




MO Graphs and Histograms
```{r MO Analysis}
MO_data <- merged_final[ merged_final$state == "MO", ]
```

```{r MO Histogram Unemployment}
hist(MO_data$value,
     main="Histogram of Unemployment Across All Years",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Missouri, from 1984 to 2018, the annual unemployment rates ranged from 2.0-9.9% The most frequent range of unemployment rates was between 5.0-5.9% which occurred 13 out of the 36 years. The next most frequent ranges were between 3.0-3.9% & 4.0-4.9% which both only appeared 6 times each.

```{r MO Histogram Poverty}
hist(MO_data$poverty,
     main="Histogram of Poverty Across All Years",
     xlab="Poverty Rate",
     breaks = 5)
```
In Missouri, from 1984-2014, the annual poverty rates ranged from 8.0-17.9%. 20 of the 36 years were characterized by poverty rates between 12.0-15.9%. Only two years displayed higher poverty rates. 

```{r MO scatter unemployment}
scatter.smooth(MO_data$year, MO_data$value, main="MO Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue", span = 2/8)
```
The annual unemployment rate in Missouri generally decreased from 1984-1999. It generally increased from 2000-2009 before sharply declining from 2010-2018. There is no observable patter or relationship based off of this plot. 
```{r MO scatter HPI}
scatter.smooth(MO_data$year, MO_data$HPI, main="MO HPI by Year", xlab = "Year", ylab = "HPI", col = "blue", span = 2/8)
```
The housing price index in Missouri consistently increased in a linear fashion from 1984 to 2008. The next several years were characterized by a steady decline until 2013 where the HPI restarted it's consistent increase year over year. 

```{r MO scatter poverty}
scatter.smooth(x=MO_data$year, y=MO_data$poverty, main="MO % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r MO scatter log RMHI}
#MO_data$log_RMHI <- as.integer(MO_data$log_RMHI)
scatter.smooth(x=MO_data$year, y=MO_data$log_RMHI, main="MO Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r MO scatter log population}
#MO_data$log_pop <- as.integer(MO_data$log_pop)
scatter.smooth(x=MO_data$year, y=MO_data$log_pop, main="MO Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r MO boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(MO_data$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(MO_data$value)$out))  # box plot for 'Unemployment'
```

```{r MO boxplot poverty}
boxplot(MO_data$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(MO_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```





WA Graphs and Histograms 
```{r WA Analysis}
WA_data <- merged_final[ merged_final$state == "WA", ]
```

```{r WA Histogram Unemployment}
hist(WA_data$value,
     main="Histogram of Unemployment Across All Years",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Washington, from 1984-2018,the annual unemployment rates ranged from 4.0-10.9%.  The histogram displays a positively skewed distribution meaning relatively low poverty rates occurred more than higher rates. There is a small presence of higher poverty rates and only 1 year that experienced a poverty rate above 10%.
```{r WA Histogram Poverty}
hist(WA_data$poverty,
     main="Histogram of Poverty Across All Years",
     xlab="Poverty Rate",
     breaks = 5)
```
In Washington, from 1984 to 2018, The annual poverty rates ranged from 8.0-12.9%. The histogram appears to be slightly negatively skewed. 18 of the 36 years were characterized by poverty rates the occurred in the two highest intervals, 11.0-11.9% and 12.0-12.9%.

```{r WA scatter unemployment}
scatter.smooth(WA_data$year, WA_data$value, main="WA Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue", span = 2/8)
```
The annual unemployment rate in Washington appears to be cyclical. The up and down pattern repeated itself 3-4 times before its intensity increased in 2008 when the unemployment rate spiked. The following years were characterized by a decline back to levels near the floor of the original pattern.   
```{r WA scatter HPI}
scatter.smooth(WA_data$year, WA_data$HPI, main="WA HPI by Year", xlab = "Year", ylab = "HPI", col = "blue", span = 2/8)
```
The housing price index in Washington consistently increased from 1984-2003 before experiencing a sharper increase from 2003-2008. After what is characteristic of a correction, between 2008-2012, the HPI began to repeat it's consistent increase year over year from 2013-2018. 
```{r WA scatter poverty}
scatter.smooth(x=WA_data$year, y=WA_data$poverty, main="WA % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r WA scatter log RMHI}
#WA_data$log_RMHI <- as.integer(WA_data$log_RMHI)
scatter.smooth(x=WA_data$year, y=WA_data$log_RMHI, main="WA Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r WA scatter log population}
#WA_data$log_pop <- as.integer(WA_data$log_pop)
scatter.smooth(x=WA_data$year, y=WA_data$log_pop, main="WA Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```


```{r WA boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(WA_data$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(WA_data$value)$out))  # box plot for 'Unemployment'
```

```{r WA boxplot poverty}
boxplot(WA_data$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(WA_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```








OR Graphs and Histograms
```{r OR Analysis}
OR_data <- merged_final[ merged_final$state == "OR", ]
```

```{r OR Histogram Unemployment}
hist(OR_data$value,
     main="Histogram of Unemployment Across All Years",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Oregon, from 1984 to 2018, the annual unemployment rates ranged from 2.0%-11.9%.The histogram did not produce a distribution with a skew worth acknowledging, however, the two most frequent ranges were 4.0-4.9% & 5.0-5.9% consecutively. There was only one year with an unemployment rate of 10.0% or higher.
```{r OR Histogram Poverty}
hist(OR_data$poverty,
     main="Histogram of Poverty Across All Years",
     xlab="Poverty Rate",
     breaks = 5)
```
In Oregon, from 1984 to 2018, the annual unemployment rates ranged from 9.0%-14.9%. The most frequent range of unemployment rates was 11.0-11.9% which occurred 15 of the 36 years.  
```{r OR scatter unemployment}
scatter.smooth(OR_data$year, OR_data$value, main="OR Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue", span = 2/8)
```
The annual unemployment rates in Oregon displayed a cyclical pattern. Like every state state observed, other than Texas, the unemployment rates peaked in 2008-2009.  The sharp decline that proceeded led to Oregon's lowest unemployment levels (2016-2018) in the 36 year time period.  
```{r OR scatter HPI}
scatter.smooth(OR_data$year, OR_data$HPI, main="OR HPI by Year", xlab = "Year", ylab = "HPI", col = "blue", span = 2/8)
```
The housing price index in Oregon consistently increased from 1984-2003 before experiencing a sharper increase from 2004-2008. The rapid increase was halted in 2008 where it began to fall until 2013 where it reached the levels of 2004. This was the new floor as the HPI began to increase in a linear fashion from 2013-2018.
```{r OR scatter poverty}
scatter.smooth(x=OR_data$year, y=OR_data$poverty, main="OR % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r OR scatter log RMHI}
#OR_data$log_RMHI <- as.integer(OR_data$log_RMHI)
scatter.smooth(x=OR_data$year, y=OR_data$log_RMHI, main="OR Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r OR scatter log population}
#OR_data$log_pop <- as.integer(OR_data$log_pop)
scatter.smooth(x=OR_data$year, y=OR_data$log_pop, main="OR Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r OR boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(OR_data$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(OR_data$value)$out))  # box plot for 'Unemployment'
```

```{r OR boxplot poverty}
boxplot(OR_data$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(OR_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```






CO Graphs and Histograms
```{r CO Analysis}
CO_data <- merged_final[ merged_final$state == "CO", ]
```

```{r CO Histogram Unemployment}
hist(CO_data$value,
     main="Histogram of Unemployment Across All Years",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Colorado, from 1984 to 2018, the annual unemployment rates ranged from 2.0% to 8.9%. 11 of the 36 years were characterized by unemployment rates between the range of 5.0-5.9%. 
```{r CO Histogram Poverty}
hist(CO_data$poverty,
     main="Histogram of Poverty Across All Years",
     xlab="Poverty Rate",
     breaks = 5)
```
In Colorado, from 1984 to 2018, the poverty rates ranged from 8.0% to 13.9%. The histogram displays a slightly positively skewed distribution. 24 of the 36 years were characterized by poverty rates in the three ranges with the lowest values, 8.0-8.9%, 9.0-9.9%, and 10.0-10.9%.  
```{r CO scatter unemployment}
scatter.smooth(CO_data$year, CO_data$value, main="CO Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue", span = 2/8)
```
The annual unemployment levels in Colorado loosely exhibits the cyclic nature seen in other states such as Oregon, Washington, and Illinois. The highest rates of unemployment occurred in 2010, a little later than the other states observed excluding Texas. The lowest unemployment rate occurred in 2000.    
```{r CO scatter HPI}
scatter.smooth(CO_data$year, CO_data$HPI, main="CO HPI by Year", xlab = "Year", ylab = "HPI", col = "blue", span = 2/8)
```
The housing price index for Colorado showed a steady increase from 1984 to 2006. The proceeding 6 years were characterized by a steady decline until 2011. From 2011-2018 the the HPI increased at a faster rate than it did from 1984-2006.  
```{r CO scatter poverty}
scatter.smooth(x=CO_data$year, y=CO_data$poverty, main="CO % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r CO scatter log RMHI}
#CO_data$log_RMHI <- as.integer(CO_data$log_RMHI)
scatter.smooth(x=CO_data$year, y=CO_data$log_RMHI, main="CO Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r CO scatter log population}
#CO_data$log_pop <- as.integer(CO_data$log_pop)
scatter.smooth(x=CO_data$year, y=CO_data$log_pop, main="CO Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r CO boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(CO_data$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(CO_data$value)$out))  # box plot for 'Unemployment'
```

```{r CO boxplot poverty}
boxplot(CO_data$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(CO_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```






WV Graphs and Histograms
```{r WV Analysis}
WV_data <- merged_final[ merged_final$state == "WV", ]
```

```{r WV Histogram Unemployment}
hist(WV_data$value,
     main="Histogram of Unemployment Across All Years",
     xlab="Unemployment Rate",
     breaks = 5)
```
In West Virginia,from 1984 to 2018, the unemployment rates ranged from 4.0-15.9%. The histogram displays a positively skewed distribution. 15 of the 36 years were characterized by unemployment rates ranging in lowest interval, 4.0-5.9%. 
```{r WV Histogram Poverty}
hist(WV_data$poverty,
     main="Histogram of Poverty Across All Years",
     xlab="Poverty Rate",
     breaks = 5)
```
In West Virginia, from 1984 to 2018, the annual poverty rates ranged from 14.0% to 23.9%. This is the highest range out of all 8 states examined. Although this range has larger values comparatively, the histogram displays a positively skewed distribution. 25 out of the 36 years were characterized by unemployment rates that fell into the first two intervals, 14.0-15.9% & 16-17.9%.  
```{r WV scatter unemployment}
scatter.smooth(WV_data$year, WV_data$value, main="WV Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue", span = 2/8)
```
The annual unemployment rate in West Virginia looks much different than the other states observed. In 1984, the unemployment rate was by far the highest during the 36 year time period. Although it did double between 2006-2010, the unemployment rate in 2010 was only a little over 8% compared to 1984 which peaked over 15%. 
```{r WV scatter HPI}
scatter.smooth(WV_data$year, WV_data$HPI, main="WV HPI by Year", xlab = "Year", ylab = "HPI", col = "blue", span = 2/8)
```
The housing price index in West Virginia is similar to the other states observed. It displayed a linear distribution, before spiking from 2003-2008. The decline from 2008-2010 was less intense and shorter lasting than many of the other states before continuing the original upward trend.
```{r WV scatter poverty}
scatter.smooth(x=WV_data$year, y=WV_data$poverty, main="WV % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r WV scatter log RMHI}
#WV_data$log_RMHI <- as.integer(WV_data$log_RMHI)
scatter.smooth(x=WV_data$year, y=WV_data$log_RMHI, main="WV Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r WV scatter log population}
#WV_data$log_pop <- as.integer(WV_data$log_pop)
scatter.smooth(x=WV_data$year, y=WV_data$log_pop, main="WV Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r WV boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(WV_data$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(WV_data$value)$out))  # box plot for 'Unemployment'
```

```{r WV boxplot poverty}
boxplot(WV_data$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(WV_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```






ME Graphs and Histograms
```{r ME Analysis}
ME_data <- merged_final[ merged_final$state == "ME", ]
```

```{r ME Histogram Unemployment}
hist(ME_data$value,
     main="Histogram of Unemployment Across All Years",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Maine, from 1984-2018, the annual unemployment rate ranged from 3.0-8.9%. The histogram displays a positively skewed distribution. 18 of the 36 years were characterized by annual unemployment rates in lowest two intervals, 3.0%-3.9% & 4.0-4.9%.
```{r ME Histogram Poverty}
hist(ME_data$poverty,
     main="Histogram of Poverty Across All Years",
     xlab="Poverty Rate",
     breaks = 5)
```
In Maine, from 1984-2018, the annual poverty rates ranged from 9.0%-15.9%. The distribution displayed by the histogram is not symmetrical, however, the interval at each end of the range experienced only one occurrence signifying outliers, to a degree. 

```{r ME scatter unemployment}
scatter.smooth(ME_data$year, ME_data$value, main="ME Unemployment by Year", xlab = "Year", ylab = "Unemployment", col = "blue", span = 2/8)
```
The annual unemployment rate in Maine displays the cyclical nature that has been observed in many of the other states. The up and down pattern floors around 3% while the peak came in 2008-2009. 
```{r ME scatter HPI}
scatter.smooth(ME_data$year, ME_data$HPI, main="ME HPI by Year", xlab = "Year", ylab = "HPI", col = "blue", span = 2/8)
```

```{r ME scatter poverty}
scatter.smooth(x=ME_data$year, y=ME_data$poverty, main="ME % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```
The housing price index in Maine increased from 1984-1989 before slightly decreasing until 1995. This was followed by a consistent increase that peaked in 2008. After the short period of decline the HPI bounced and continued to rise until the end of the time period in 2018, like many of the states examined. 

```{r ME scatter log RMHI}
#ME_data$log_RMHI <- as.integer(ME_data$log_RMHI)
scatter.smooth(x=ME_data$year, y=ME_data$log_RMHI, main="ME Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r ME scatter log population}
#ME_data$log_pop <- as.integer(ME_data$log_pop)
scatter.smooth(x=ME_data$year, y=ME_data$log_pop, main="ME Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r ME boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(ME_data$value, main="Unemployment", sub=paste("Outlier rows: ", boxplot.stats(ME_data$value)$out))  # box plot for 'Unemployment'
```

```{r ME boxplot poverty}
boxplot(ME_data$poverty, main="Poverty", sub=paste("Outlier rows: ", boxplot.stats(ME_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```