---
title: "Fred"
author: "https://github.com/jrkkv5/FIN6572Fall2020/blob/master/National%20Debt/importNationalDebt.Rmd"
date: "11/8/2020"
output: html_document
---

```{r library}
#Loading necessary packages
library(readr)
library(devtools)
library(blsAPI)
library(rjson)
library(curl)
library(dplyr)
library(tidyverse)
library(rmarkdown)
library(knitr)
library(pastecs)
library(ggplot2)
library(corrplot)
library(lmtest)
library(RCurl)
library(forecast)
library(olsrr)

#***ADD LIBRARY(olsrr) ******
#Update code for your states and run regression from 1984-2014, create some basic plots, etc.
```

```{r US Descriptive Statistics}
#Descriptive Statistics
stat.desc(merged_final)
```

The dependent variable, value, has a mean of 5.44 and a standard deviation of 1.97. Connecticut and Virginia had the lowest unemployment rate in 2000 at 1.8, while West Virginia was the highest in 1984 at 14.4. The mean HPI was 260.15, with a standard deviation of 119.68. The lowest HPI was found in Wyoming in 1987 at 79.25, while the highest was in Massachusetts in 2018 at 796.86. Poverty rate had a mean of 12.85 and a standard deviation of 3.69. In 1989, Connecticut had the lowest rate of 2.9, and Mississippi had the highest at 27.2 in 1988. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 15.06, and the standard deviation was 1.01. Wyoming had the lowest logged population at 13.025 in 1990, California having the highest in 2018 at 17.49. Logged real median household income had a mean of 10.95 and a standard deviation of 0.17. In 2013, Mississippi had the lowest log_RMHI at 10.46, while Massachusetts had the highest in 2018 at 11.37. 


This project uses the ordinary least squares method to estimate the relationship between the independent variable and the dependent variable. The following model was estimated using OLS:
```{r US multiple linear regression log all}
US_reg1 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI , data = merged_final)
summary(US_reg1)
anova(US_reg1)
ols_coll_diag(US_reg1)
```

The first multiple regression for US is conducted with five independent variables, four of the coefficients on the predictor variables have the expected signs except the logged real median household income. Logged real median household income has a positive sign when it in fact should have a negative one. The wrong sign could be explained partly by it having the highest VIF value of 3.831, but it was not high enough to indicate multi-collinearity. 
According to the low VIF values between 1.029 and 3.831 the variables are not correlated with each other. The low VIF values make the p-value reliable to test significance of the independent variables. The p values indicate that three of the variables are statistically significant at the 5% level of significance except HPI and sp500. A second regression will be made excluding the sp500 to improve the model.
The overall fit of the estimated model does not appear to be good because the R-sq of the regression is low. The models R-sq is 0.2511 and the adjusted R-sq is 0.249 which indicates that the model explains 25.11% of the variation in the unemployment rate data around its mean.

```{r US multiple linear significant variables}
US_linearModelSignificant <- lm(value ~ poverty + log_pop + log_RMHI , data = merged_final)
summary(US_linearModelSignificant)
anova(US_linearModelSignificant)
ols_coll_diag(US_linearModelSignificant)
```

The second regression result show that log_RHMI still has the incorrect sign. With the two variables excluded, the VIF values decrease ranging between 1.160 and 3.027, indicating no multi-collinearity among the independent variables. The p-values were still statistically significant at the 5% level of significance. Log_pop is the least significant variable, but it has the lowest VIF value. 
The overall fit of the estimated model does not appear to be good because the R-sq of the regression is low. The models R-sq is 0.2493 and the adjusted R-sq is 0.2481 which decreased compared to the original regression.

```{r IL Descriptive Statistics}
#Descriptive statistics
stat.desc(IL_data)
```

The dependent variable, unemployment, has a mean of 6.44 and a standard deviation of 1.81. The lowest unemployment rate was in 1998 and 1999 at 4.1, while the highest was in 2009 at 11.00. The mean HPI was 256.07, with a standard deviation of 79.06. The lowest HPI was in 1984 at 112.80, while the highest was in 2018 at 371.21. Poverty rate had a mean of 12.54 and a standard deviation of 1.58. In 1999, lowest rate was seen of 9.9, and the highest was 15.6 in 1992. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 16.32, and the standard deviation was 0.046. The lowest logged population at 16.248 in 1986, and the highest in 2013 at 16.373. Logged real median household income had a mean of 11.02 and a standard deviation of 0.066. In 1984, Illinois had the lowest log_RMHI at 10.91, while had the highest in 2018 at 11.16. 


This project uses the ordinary least squares method to estimate the relationship between the independent variable and the dependent variable. The following model was estimated using OLS:
```{r IL multiple linear regression log all}
IL_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI , data = IL_data)
summary(IL_reg2)
anova(IL_reg2)
ols_coll_diag(IL_reg2)
```

The first multiple regression for Illinois is conducted with five independent variables, four of the coefficients on the predictor variables have the expected signs except the sp500. Sp500 has a positive sign when it in fact should have a negative one. The wrong sign could be explained partly by the test high VIF values indicating some variables have high multi-collinearity within the model. 
According to the high VIF values between 1.080 and 9.367, the variables are correlated with each other. The high VIF values make the p-value not reliable to test significance of the independent variables. A second regression is made excluding the HPI which has the highest VIF value to improve the model.
The overall fit of the estimated model appears to be good because the R-sq of the regression is in the mid-level. The models R-sq is 0.6109 and the adjusted R-sq is 0.546 which indicates that the model explains 61.09% of the variation in the unemployment rate data around its mean.

```{r IL multiple linear significant variables}
IL_linearModelSignificant <- lm(value ~ poverty + sp500 + log_pop + log_RMHI, data = IL_data)
summary(IL_linearModelSignificant)
anova (IL_linearModelSignificant)
ols_coll_diag(IL_linearModelSignificant)

```

The second regression result show that sp500 still has the incorrect sign. With the two variables excluded, the VIF values significantly decrease ranging between 1.030 and 4.417, indicating no multi-collinearity among the independent variables. The p-values are statistically significant at the 5% level of significance for poverty and log_pop. Sp500 and Log_RMHI are not statistically significant according to their p-values. 
The overall fit of the estimated model appears to be good because the R-sq of the regression is in the mid-level. The models R-sq is 0.6066 which decreased compared to the first regression, but the adjusted R-sq, 0.5559, increased compared to the first regression.

```{r WV Descriptive Statistics}
#Descriptive statistics
stat.desc(WV_data)
```

The dependent variable, unemployment, has a mean of 7.25 and a standard deviation of 2.48. The lowest unemployment rate was in 2006 at 4.2 while the highest was in 1984 at 14.40. The mean HPI was 164.11, with a standard deviation of 48.67. The lowest HPI was in 1984 at 87.28, while the highest was in 2018 at 230.86. Poverty rate had a mean of 17.60 and a standard deviation of 2.42. In 2004, lowest rate was seen of 14.2, and the highest was 22.4 in 1986. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 14.42, and the standard deviation was 0.016. The lowest logged population at 14.40 in 1990, and the highest in 1984 at 14.47. Logged real median household income had a mean of 10.67 and a standard deviation of 0.10. In 1992, West Virginia had the lowest log_RMHI at 10.48, while had the highest in 2007 at 10.84.


This project uses the ordinary least squares method to estimate the relationship between the independent variable and the dependent variable. The following model was estimated using OLS:
```{r WV multiple linear regression log all}
WV_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI, data = WV_data)
summary(WV_reg2)
anova(WV_reg2)
ols_coll_diag(WV_reg2)
```

The first multiple regression for West Virginia is conducted with five independent variables, three of the coefficients on the predictor variables have the expected signs except sp500 and log_RMHI. Log_RMHI and sp500 have a positive slope when it in fact they should have a negative one. The wrong signs could be explained partly by the test high VIF values indicating some variables have high multi-collinearity within the model.
According to the high VIF values between 1.271 and 8.043, the variables are correlated with each other. Only one of the variables, Log_RMHI, has a VIF value close to the critical limit of 10, so the p-values are used to test significance of the independent variables. The p values indicate that four of the variables are statistically significant at the 5% level of significance except sp500. A second regression is made excluding the sp500 to improve the model.
The overall fit of the estimated model appears to be substantial because the R-sq of the regression is high. The models R-sq is 0.801 and the adjusted R-sq is 0.7678 which indicates that the model explains 80.10% of the variation in the unemployment rate data around its mean.

```{r WV multiple linear significant variables}
WV_linearModelSignificant <- lm(value ~ HPI + poverty + log_pop + log_RMHI, data = WV_data)
summary(WV_linearModelSignificant)
anova(WV_linearModelSignificant)
ols_coll_diag(WV_linearModelSignificant)
```

The second regression result show that log_RHMI still has the incorrect sign. With the two variables excluded, the VIF values decrease ranging between 1.279 and 6.623, but HPI and log_RMHI still show signs of multi-collinearity. The p-values are still statistically significant at the 5% level of significance.
The overall fit of the estimated model appears to be substantial because the R-sq of the regression is high. The models R-sq is 0.7951 which decreased compared to the first regression, but the adjusted R-sq, 0.7687, increased compared to the first regression

#Conclusion of the Multiple Regressions results
Several conclusions were made using the multiple regression models. In all the models tested, sp500 is never significant showing that the index is less of a factor in determining the unemployment. The models show that unemployment rate might not be the sole factor in indicating the economy overall health. Poverty and logged population are significant on all the models and have effect on unemployment rate with the expected sign. There was a significance effect on unemployment rate by logged real median household income on all the models with reliable p-values, but it did not have the expected sign. The modes show that an increase in real median household income would increase unemployment rate which will be contrary to the initial hypothesis. The relationship between house price index and unemployment is not consistent within all the test models. 