---
title: "Unemployment Report"
author: "Jeff Krueger, Corey Lund, Fred Mwangi, and Nemanja Orescanin"
date: "11/29/2020"
output:
<<<<<<< HEAD
  html_document: default
  pdf_document: default
=======
  word_document: default
  html_document: default
>>>>>>> 32a2b1c8121c34957d746e37902519e2611353fb
---

```{r packages}
library(devtools)
library(blsAPI)
library(rjson)
library(curl)
library(RCurl)
library(knitr)
library(readr)
library(dplyr)
library(tidyverse)
library(rmarkdown)
library(pastecs)
library(ggplot2)
library(corrplot)
library(lmtest)
library(forecast)
#loads the necessary packages

```

First we need to load the libraries that we will be using into our report. The above code should load the proper libraries. R Studio will prompt you to install the packages if necessary.


``` {r log variables}
merged_final <- read_csv("csv files/mergedFinal.csv")
#Reads final file into merged dataframe

merged_final$log_pop = as.numeric(log(merged_final$population))

merged_final$log_RMHI = as.numeric(log(merged_final$RMHI))
```
As a final step in scrubbing our data, we created new variables to log the population and median household income variables and store it in our dataframe. By taking the logarithmic values, these variables become more comparable to the existing variables we have within our dataset.

Below you can see the table output once all of our data has been merged.
``` {r merged_final_table}

kable(merged_final[1:6,], caption = "Table Including All Variables")
```

```{r All data Sescriptive Statistics}
#Descriptive Statistics
stat.desc(merged_final)
```

The dependent variable, value, has a mean of 5.44 and a standard deviation of 1.97. Connecticut and Virginia had the lowest unemployment rate in 2000 at 1.8, while West Virginia was the highest in 1984 at 14.4. The mean HPI was 260.15, with a standard deviation of 119.68. The lowest HPI was found in Wyoming in 1987 at 79.25, while the highest was in Massachusetts in 2018 at 796.86. Poverty rate had a mean of 12.85 and a standard deviation of 3.69. In 1989, Connecticut had the lowest rate of 2.9, and Mississippi had the highest at 27.2 in 1988. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 15.06, and the standard deviation was 1.01. Wyoming had the lowest logged population at 13.025 in 1990, California having the highest in 2018 at 17.49. Logged real median household income had a mean of 10.95 and a standard deviation of 0.17. In 2013, Mississippi had the lowest log_RMHI at 10.46, while Massachusetts had the highest in 2018 at 11.37. 

After our data was merged we binned the states into groups based on the mean population of the years in our dataframe (1984-2018).  Below are the bins that were created:

Population < 2,000,000
WV, NM, NE, ID, ME, HI, NH, RI, MT, DE, SD, ND, AK, VT, WY

2,000,000 < Population < 5,000,000
MN, AL, LA, CO, SC, KY, OK, OR, CT, IA, MS, KS, AR, UT, NV

5,000,000 < Population < 10,000,000
MI, NJ, GA, NC, VA, MA, IN, WA, TN, MO, WI, MD, AZ

Population > 10,000,000
CA, TX, NY, FL, PA, IL, OH


Each member of our group performed some basic analysis on the states within one of the bins. After performing our initial analysis we chose to focus on 2 states within each bin.  The states we chose to focus on are: TX, IL, MO, WA, OR, CO, WV, ME


Below are some of the highlights along with commentary for the states we have chosen to focus on 

```{r IL Analysis}
IL_data <- merged_final[ merged_final$state == "IL", ]
```

```{r IL Descriptive Statistics}
#Descriptive statistics
stat.desc(IL_data)
```

The dependent variable, unemployment, has a mean of 6.44 and a standard deviation of 1.81. The lowest unemployment rate was in 1998 and 1999 at 4.1, while the highest was in 2009 at 11.00. The mean HPI was 256.07, with a standard deviation of 79.06. The lowest HPI was in 1984 at 112.80, while the highest was in 2018 at 371.21. Poverty rate had a mean of 12.54 and a standard deviation of 1.58. In 1999, lowest rate was seen of 9.9, and the highest was 15.6 in 1992. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 16.32, and the standard deviation was 0.046. The lowest logged population at 16.248 in 1986, and the highest in 2013 at 16.373. Logged real median household income had a mean of 11.02 and a standard deviation of 0.066. In 1984, Illinois had the lowest log_RMHI at 10.91, while had the highest in 2018 at 11.16. 

```{r IL Pairwise Correlations}
#pairwise correlations
sapply(IL_data, class)
sapply(IL_data, is.factor)
cor(IL_data[sapply(IL_data, function(x) !is.character(x))])
```

```{r IL Histogram Unemployment}
hist(IL_data$value,
     main="Histogram of Unemployment from 1984-2018",
     xlim=c(4,11),
     xlab="Unemployment Rate",
     breaks = 5)
```

Histograms graphically summarize the distribution of a data set. In Illinois, from 1984 to 2018, The annual unemployment rates ranged from 4.0-11.0%. The histogram displays a positively skewed distribution. An unemployment rate that fell within the first two intervals, 4.0-4.9% & 5.0-5.9%, occurred 18 out of the 36 years examined.  

```{r IL Histogram Poverty}
hist(IL_data$poverty,
     main="Histogram of Poverty from 1984-2018",
     xlim=c(9,16),
     xlab="Poverty Rate",
     breaks = 5)
```

In Illinois, from 1984 to 2018, the poverty levels ranged from 9.0-16.0%. The histogram displays a relatively normal distribution. Between 12.0-12.9% was the most frequently recorded interval during this time period. In 14 of the 36 years examined, the poverty rate in Illinois was between 13-15.9%. 

```{r IL scatter unemployment}
scatter.smooth(IL_data$year, IL_data$value, main="IL Unemployment by Year", xlab = "Year", ylab = "Unemployment", span = 2/8)
```

A scatterplot displays the relationship between two numeric variables. This plot demonstrates the relationship between the annual unemployment rate and the year, from 1984 to 2018. In Illinois, there is no linear relationship, however, there is an observable cyclical pattern. The annual unemployment rate was lowest around 1999 and peaked in 2008. 

```{r IL scatter HPI}
scatter.smooth(IL_data$year, IL_data$HPI, main="IL HPI by Year", xlab = "Year", ylab = "HPI", span = 2/8)
```

This plot represents the relationship between Illinois' median housing price index (HPI) and the year. From 1984 to 2008 the relationship appeared to be linear until the HPI dipped in 2008-2013. Since then, the HPI began to continue its increase in a linear fashion. This pattern is observed consistently when examining many other states and, through inference, can partially be attributed to the housing crisis that occurred in 2008 and 2009.

```{r IL scatter poverty}
scatter.smooth(x=IL_data$year, y=IL_data$poverty, main="IL % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

The percentage of individuals living below the poverty rate in Illinois does not display a clear pattern. Overall, there has been slight decrease of roughly 4% from 1984 to 2018. Poverty levels peaked in 1992.

```{r IL scatter log RMHI}
#IL_data$log_RMHI <- as.integer(IL_data$log_RMHI)
scatter.smooth(x=IL_data$year, y=IL_data$log_RMHI, main="IL Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

The median household income in Illinois increased dramatically from 1984-2018. The lowest and highest values recorded fell at the beginning and the end of the time period, respectively.

```{r IL scatter log population}
#IL_data$log_pop <- as.integer(IL_data$log_pop)
scatter.smooth(x=IL_data$year, y=IL_data$log_pop, main="IL Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

The log population of Illinois decreased from 1984-1988 before increasing in a relatively linear fashion from 1988 to its peak in 2013. 2013-2018 is characterized by a gradual decline in population. 

```{r IL correlations}
cor(IL_data$RMHI, IL_data$value)
cor(IL_data$HPI, IL_data$value)
cor(IL_data$poverty, IL_data$value)
cor(IL_data$population, IL_data$value)
cor(IL_data$sp500, IL_data$value)
cor(IL_data$log_pop, IL_data$value)
cor(IL_data$log_RMHI, IL_data$value)
#run base correlations between categories and dependent variable
```

A correlation function is used to illustrate the relationship between our independent variable and our dependent variable. The closer the correlation is to 1 or -1 the stronger the correlation between the two variables. For a correlation of 1, it signifies that there is a perfect positive correlation. For a -1 correlation there is a perfect inverse relationship between the two variables. 

The correlation between the RMHI and our dependent variable, unemployment, in Illinois is 0.03109792. The correlation shows that there is a weak positive correlation between median household income 

The correlation between HPI (Housing Price Index) and our dependent variable, unemployment, in Illinois is -0.08115831. The correlation shows that there is a weak negative relationship between the housing price index and unemployment. 

The correlation between poverty, the percentage of the population under the poverty line, and unemployment in Illinois is 0.6975789. This correlation shows a much stronger positive relationship between the two variables. As unemployment increases there is a much stronger possibility that the percentage of the population under the poverty line increasing as well. 

The correlation between the population and unemployment in Illinois is 0.02011968. The correlation between population and unemployment is a weak positive correlation. The correlation is extremely close to 0, which signified that there is no direct correlation between the two variables. 

The correlation between the s&p 500 and unemployment in Illinois is 0.03109792. The correlation between the S&P 500 and unemployment is a weak positive correlation. With the correlation being extremely close to 0 it signifies that there is not much of a correlation between the two variables. 

The correlation between the log of the population and unemployment in Illinois is 0.01123117. The correlation between the log of the population and unemployment is extremely close to 0, which signifies there is almost no correlation between the two variables.  

The correlation between the log of median household income and unemployment in Illinois is -0.6602776. The correlation between the log of median household income and unemployment is a strong negative correlation. A strong negative correlation suggests that the two variables move in inverse directions of each other. In this case, -0.66 suggests that there is enough evidence to suggest that in most cases the two variables move in inverse directions of each other due to the proximity of -0.66 to -1 which is a perfect negative relationship.
```{r IL boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(IL_data$value, main="Unemployment in IL", sub=paste("Outlier rows: ", boxplot.stats(IL_data$value)$out))  # box plot for 'Unemployment'
```

The annual unemployment rate in Illinois ranged from 4%-11% from 1984-2018. The boxplot displays a positively skewed distribution meaning the mean is greater than the median. There were no outliers in the Unemployment data collected from 1984-2018 based upon the boxplot results.

```{r IL boxplot poverty}
boxplot(IL_data$poverty, main="IL Poverty", sub=paste("Outlier rows: ", boxplot.stats(IL_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```

The poverty levels in Illinois ranged from roughly 9% to just under 14% from 1984-2018. The median poverty rate is displayed at roughly 12.5%.The plot displays a relatively normal distribution meaning the mean and median are close the same value. There were no outliers in the Poverty data collected from 1984-2018 based upon the boxplot results.

```{r IL linear regression all}
#run multiple linear model for data
IL_reg1 <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = IL_data)
summary(IL_reg1)
anova(IL_reg1)
```

```{r IL linear regression log all}
IL_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI , data = IL_data)
summary(IL_reg2)
anova(IL_reg2)
```

```{r IL linear significant variables}
IL_linearModelSignificant <- lm(value ~ poverty + log_pop, data = IL_data)
summary(IL_linearModelSignificant)
```



<<<<<<< HEAD
=======

```{r TX Analysis}
TX_data <- merged_final[ merged_final$state == "TX", ]
```

```{r TX Descriptive Statistics}
#Descriptive statistics
stat.desc(TX_data)
```

The dependent variable, unemployment, has a mean of 5.70 and a standard deviation of 1.35. The lowest unemployment rate was in 2000 at 3.5, while the highest was in 1986 at 8.5. The mean HPI was 185.77, with a standard deviation of 61.34. The lowest HPI was in 1988 at 113.77, while the highest was in 2018 at 334.92. Poverty rate had a mean of 16.46 and a standard deviation of 1.37. In 2017, lowest rate was seen of 13.2 and the highest was 19.1 in 1994. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 16.88, and the standard deviation was 0.17. The lowest logged population at 16.59 in 1984, and the highest in 2018 at 17.17. Logged real median household income had a mean of 10.91 and a standard deviation of 0.056. In 1992, Texas had the lowest log_RMHI at 10.80, while had the highest in 2017 at 11.03. 

```{r TX Pairwise Correlations}
#pairwise correlations
sapply(TX_data, class)
sapply(TX_data, is.factor)
cor(TX_data[sapply(TX_data, function(x) !is.character(x))])
```


```{r TX Histogram Unemployment}
hist(TX_data$value,
     main="Histogram of Unemployment from 1984-2018",
     xlim=c(3,9),
     xlab="Unemployment Rate",
     breaks = 5)
```
In Texas, from 1984 to 2018, the annual unemployment rates ranged from 3.0-8.9%. The most common interval, or range of annual unemployment rates was between 5.0-5.9% which occurred 10 times in the 36 year period. Texas only experienced one year with an unemployment rate of 8.0% or higher. 

```{r TX Histogram Poverty}
hist(TX_data$poverty,
     main="Histogram of Poverty from 1984-2018",
     xlim=c(13,20),
     xlab="Poverty Rate",
     breaks = 5)
```
In Texas, from 1984 to 2018, the annual poverty rates ranged from 13.0-20.0%. The histogram is nearly symmetrical meaning most of the values occurred in the near middle of the range. 27 of the 36 years were characterized by poverty rates ranging from 15.0-17.9%. Only 5 years experienced lower rates and only 4 experienced higher rates of poverty.  

```{r TX scatter unemployment}
scatter.smooth(TX_data$year, TX_data$value, main="TX Unemployment by Year", xlab = "Year", ylab = "Unemployment", span = 2/8)
```

```{r TX scatter HPI}
scatter.smooth(TX_data$year, TX_data$HPI, main="TX HPI by Year", xlab = "Year", ylab = "HPI", span = 2/8)
```

```{r TX scatter poverty}
scatter.smooth(x=TX_data$year, y=TX_data$poverty, main="TX % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r TX scatter RMHI}
TX_data$RMHI <- as.integer(TX_data$RMHI)
scatter.smooth(x=TX_data$year, y=TX_data$RMHI, main="TX Median Household Income by Year", xlab = "Year", ylab = "RMHI", span = 2/8 )
```

```{r TX scatter population}
TX_data$population <- as.integer(TX_data$population)
scatter.smooth(x=TX_data$year, y=TX_data$population, main="TX Population by Year", xlab = "Year", ylab = "Population", span = 2/8 )
```

```{r TX scatter log RMHI}
#TX_data$log_RMHI <- as.integer(TX_data$log_RMHI)
scatter.smooth(x=TX_data$year, y=TX_data$log_RMHI, main="TX Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r TX scatter log population}
#TX_data$log_pop <- as.integer(TX_data$log_pop)
scatter.smooth(x=TX_data$year, y=TX_data$log_pop, main="TX Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r TX correlations}
cor(TX_data$RMHI, TX_data$value)
cor(TX_data$HPI, TX_data$value)
cor(TX_data$poverty, TX_data$value)
cor(TX_data$population, TX_data$value)
cor(TX_data$sp500, TX_data$value)
cor(TX_data$log_pop, TX_data$value)
cor(TX_data$log_RMHI, TX_data$value)
#run base correlations between categories and dependent variable
```
The correlation between the RMHI and our dependent variable, unemployment, in Texas is -0.5663074. The correlation between median household income and unemployment is a decently strong negative correlation. This is much stronger correlation than was the vase with the previous state of Illinois.

The correlation between the housing price index and unemployment in Texas is -0.4234252 which is a negative correlation. It illustrates a correlation that points towards an inverse relationship between the housing price index and unemployment. 

The correlation between poverty, the proportion of the population under the poverty line, and unemployment in Texas is 0.6719098. The correlation between poverty and unemployment shows a decently strong correlation between the two variables. The positive correlation shows that the variables move in the same direction dependent on the value of the variables.  

The correlation between population and unemployment in Texas is -0.3861266. The correlation between population and unemployment points to a weak inverse relationship. This means that as population goes up, then the unemployment tends to go down. The correlation is not as close to -1 as it is to 0, which shows that there are instances where there is no direct relationship between the two variables. 

The correlation between the S&P 500 and unemployment in Texas is 0.09162901. The correlation between S&P 500 and unemployment is a weak positive correlation. With the correlation being this close to zero it shows that in most cases there is no direct correlation between the two variables. 

The correlation between the log of the population and unemployment in Texas is -0.3941111. The correlation between the log of population and unemployment shows that there is an inverse relationship between the two variables. THe correlation is not the strongest, but does illustrate that in certain cases the two variables do have an inverse relationship. 

The correlation between the log of the median household income and unemployment in Texas is -0.5626285. THe correlation between the log of the median household income and unemployment shows a strong inverse relationship. An inverse relationship suggests that the two variables tend to go in opposite directions of each other. So, for instance, as median household income goes up,then the unemployment tends to go down.   
```{r TX boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(TX_data$value, main="TX Unemployment", sub=paste("Outlier rows: ", boxplot.stats(TX_data$value)$out))  # box plot for 'Unemployment'
```

```{r TX boxplot poverty}
boxplot(TX_data$poverty, main="TX Poverty", sub=paste("Outlier rows: ", boxplot.stats(TX_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```

```{r TX linear regression all}
#run multiple linear model for data
TX_reg1 <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = TX_data)
summary(TX_reg1)
anova(TX_reg1)
```

```{r TX linear regression log all}
TX_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI, data = TX_data)
summary(TX_reg2)
anova(TX_reg2)
```

```{r TX linear significant variables}
TX_linearModelSignificant <- lm(value ~ poverty + log_pop, data = TX_data)
summary(TX_linearModelSignificant)

```

```{r MO Analysis}
MO_data <- merged_final[ merged_final$state == "MO", ]
```

```{r MO Descriptive Statistics}
#Descriptive statistics
stat.desc(MO_data)
```

The dependent variable, unemployment, has a mean of 5.40 and a standard deviation of 1.63. The lowest unemployment rate was in 1999 at 2.8, while the highest was in 2009 at 9.4. The mean HPI was 223.98, with a standard deviation of 65.74. The lowest HPI was in 1984 at 118.94, while the highest was in 2018 at 334.60. Poverty rate had a mean of 12.81 and a standard deviation of 2.27. In 2000, lowest rate was seen of 9.2, and the highest was 17.5 in 2013. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 15.54, and the standard deviation was 0.070. The lowest logged population at 15.42 in 1984, and the highest in 2018 at 15.63. Logged real median household income had a mean of 10.91 and a standard deviation of 0.090. In 1984, Missouri had the lowest log_RMHI at 10.78, while had the highest in 2000 at 11.10. 

```{r MO Pairwise Correlations}
#pairwise correlations
sapply(MO_data, class)
sapply(MO_data, is.factor)
cor(MO_data[sapply(MO_data, function(x) !is.character(x))])
```


```{r MO Histogram Unemployment}
hist(MO_data$value,
     main="Histogram of Unemployment from 1984-2018",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Missouri, from 1984 to 2018, the annual unemployment rates ranged from 2.0-9.9% The most frequent range of unemployment rates was between 5.0-5.9% which occurred 13 out of the 36 years. The next most frequent ranges were between 3.0-3.9% & 4.0-4.9% which both only appeared 6 times each.

```{r MO Histogram Poverty}
hist(MO_data$poverty,
     main="Histogram of Poverty from 1984-2018",
     xlab="Poverty Rate",
     breaks = 5)
```
In Missouri, from 1984-2014, the annual poverty rates ranged from 8.0-17.9%. 20 of the 36 years were characterized by poverty rates between 12.0-15.9%. Only two years displayed higher poverty rates.

```{r MO scatter unemployment}
scatter.smooth(MO_data$year, MO_data$value, main="MO Unemployment by Year", xlab = "Year", ylab = "Unemployment", span = 2/8)
```

```{r MO scatter HPI}
scatter.smooth(MO_data$year, MO_data$HPI, main="MO HPI by Year", xlab = "Year", ylab = "HPI", span = 2/8)
```

```{r MO scatter poverty}
scatter.smooth(x=MO_data$year, y=MO_data$poverty, main="MO % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r MO scatter RMHI}
MO_data$RMHI <- as.integer(MO_data$RMHI)
scatter.smooth(x=MO_data$year, y=MO_data$RMHI, main="MO Median Household Income by Year", xlab = "Year", ylab = "RMHI", span = 2/8 )
```

```{r MO scatter population}
MO_data$population <- as.integer(MO_data$population)
scatter.smooth(x=MO_data$year, y=MO_data$population, main="MO Population by Year", xlab = "Year", ylab = "Population", span = 2/8 )
```

```{r MO scatter log RMHI}
#MO_data$log_RMHI <- as.integer(MO_data$log_RMHI)
scatter.smooth(x=MO_data$year, y=MO_data$log_RMHI, main="MO Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r MO scatter log population}
#MO_data$log_pop <- as.integer(MO_data$log_pop)
scatter.smooth(x=MO_data$year, y=MO_data$log_pop, main="MO Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r MO correlations}
cor(MO_data$RMHI, MO_data$value)
cor(MO_data$HPI, MO_data$value)
cor(MO_data$poverty, MO_data$value)
cor(MO_data$population, MO_data$value)
cor(MO_data$sp500, MO_data$value)
cor(MO_data$log_pop, MO_data$value)
cor(MO_data$log_RMHI, MO_data$value)
#run base correlations between categories and dependent variable
```

THe correlation between the median household income and unemployment in Missouri is -0.4832073. The correlation between the median household income and unemployment is that of a medium strength negative relationship. A negative relationship between the two variables suggests that the two variables move inversely of one another with a medium stength relationship.

The correlation between the housing price index and unemployment in Missouri is 0.05064891. The correlation between the housing price index and unemployment is that of a small positive relationship. The relationship between the two variables is quite small with it being so close to 0, but does suggest that the two variables do move in the same direction of one another. 

The correlation between poverty, the proportion of the population below the poverty line, and unemployment in Missouri is 0.603927. The correlation between poverty and unemployment is that of a large positive correlation. A large positive correlation suggests that the two variables tend to mimic each other. As poverty increases so does the unemployment rate.

The correlation between the population and unemployment in Missouri is 0.08687034. The correlation between the population and unemployment is a small positive correlation. The small positive correlation suggests a weak relationship between the two variables, but does suggest that they move in the same direction of one another. 

The correlation between the S&P 500 and unemployment in Missouri is -0.09530716. The correlation between the S&P 500 and unemployment is that of a small negative correlation. THe relationship between the two variables is a weak relationship, but does suggest that they move inversely of one another. 

The correlation of the log of the population and unemployment in Missouri is 0.08057646. The correlation between the log of the population and unemployment is that of a small positive relationship. The stregth of the correlation is quite weak, but does suggest that the two variables move in the same direction of one another.

The correlation of the log of the median household income and unemployment in Missouri is -0.4725887. The correlation between the log of the median household income and unemployment suggest a medium positive correlation. The relationship between the two variables is a medium positive relationship suggesting the two variables tend to move in the same direction of one another. 

```{r MO boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(MO_data$value, main="MO Unemployment", sub=paste("Outlier rows: ", boxplot.stats(MO_data$value)$out))  # box plot for 'Unemployment'
```

```{r MO boxplot poverty}
boxplot(MO_data$poverty, main="MO Poverty", sub=paste("Outlier rows: ", boxplot.stats(MO_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```

```{r MO linear regression all}
#run multiple linear model for data
MO_reg1 <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = MO_data)
summary(MO_reg1)
anova(MO_reg1)
```

```{r MO linear regression log all}
MO_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI, data = MO_data)
summary(MO_reg2)
anova(MO_reg2)
```

```{r MO linear significant variables}
MO_linearModelSignificant <- lm(value ~ poverty + log_pop, data = MO_data)
summary(MO_linearModelSignificant)
```

```{r WA Analysis}
WA_data <- merged_final[ merged_final$state == "WA", ]
```

```{r WA Descriptive Statistics}
#Descriptive statistics
stat.desc(WA_data)
```

The dependent variable, unemployment, has a mean of 6.54 and a standard deviation of 1.59. The lowest unemployment rate was in 1997 at 4.5, while the highest was in 2009 at 10.5. The mean HPI was 317.69, with a standard deviation of 145.17. The lowest HPI was in 1984 at 110.19, while the highest was in 2018 at 632.92. Poverty rate had a mean of 10.84 and a standard deviation of 1.30. In 2006, lowest rate was seen of 8.0, and the highest was 12.9 in 1986. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 15.59, and the standard deviation was 0.16. The lowest logged population at 15.28 in 1984, and the highest in 2018 at 15.84. Logged real median household income had a mean of 11.08 and a standard deviation of 0.088. In 1985, Washington had the lowest log_RMHI at 10.89, while had the highest in 2018 at 11.29. 

```{r WA Pairwise Correlations}
#pairwise correlations
sapply(WA_data, class)
sapply(WA_data, is.factor)
cor(WA_data[sapply(WA_data, function(x) !is.character(x))])
```

```{r WA Histogram Unemployment}
hist(WA_data$value,
     main="Histogram of Unemployment from 1984-2018",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Washington, from 1984-2018,the annual unemployment rates ranged from 4.0-10.9%.  The histogram displays a positively skewed distribution meaning relatively low poverty rates occurred more than higher rates. There is a small presence of higher poverty rates and only 1 year that experienced a poverty rate above 10%.

```{r WA Histogram Poverty}
hist(WA_data$poverty,
     main="Histogram of Poverty from 1984-2018",
     xlab="Poverty Rate",
     breaks = 5)
```
In Washington, from 1984 to 2018, The annual poverty rates ranged from 8.0-12.9%. The histogram appears to be slightly negatively skewed. 18 of the 36 years were characterized by poverty rates the occurred in the two highest intervals, 11.0-11.9% and 12.0-12.9%.

```{r WA scatter unemployment}
scatter.smooth(WA_data$year, WA_data$value, main="WA Unemployment by Year", xlab = "Year", ylab = "Unemployment", span = 2/8)
```

```{r WA scatter HPI}
scatter.smooth(WA_data$year, WA_data$HPI, main="WA HPI by Year", xlab = "Year", ylab = "HPI", span = 2/8)
```

```{r WA scatter poverty}
scatter.smooth(x=WA_data$year, y=WA_data$poverty, main="WA % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r WA scatter RMHI}
WA_data$RMHI <- as.integer(WA_data$RMHI)
scatter.smooth(x=WA_data$year, y=WA_data$RMHI, main="WA Median Household Income by Year", xlab = "Year", ylab = "RMHI", span = 2/8 )
```

```{r WA scatter population}
WA_data$population <- as.integer(WA_data$population)
scatter.smooth(x=WA_data$year, y=WA_data$population, main="WA Population by Year", xlab = "Year", ylab = "Population", span = 2/8 )
```

```{r WA scatter log RMHI}
#WA_data$log_RMHI <- as.integer(WA_data$log_RMHI)
scatter.smooth(x=WA_data$year, y=WA_data$log_RMHI, main="WA Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r WA scatter log population}
#WA_data$log_pop <- as.integer(WA_data$log_pop)
scatter.smooth(x=WA_data$year, y=WA_data$log_pop, main="WA Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r WA correlations}
cor(WA_data$RMHI, WA_data$value)
cor(WA_data$HPI, WA_data$value)
cor(WA_data$poverty, WA_data$value)
cor(WA_data$population, WA_data$value)
cor(WA_data$sp500, WA_data$value)
cor(WA_data$log_pop, WA_data$value)
cor(WA_data$log_RMHI, WA_data$value)
#run base correlations between categories and dependent variable
```
The correlation between the median household income and unemployment in Washington is -0.381975. The correlation between the median household income and unemployment is that of a medium negative correlation. The relationship between the two variables has medium strength and suggests that they move inversely of one another. 

The correlation between the housing price index and unemployment in Washington is -0.1978439. The correlation between the housing price index and unemployment is that of a small negative correlation. A small negative correlation between the two variables suggests a weak inverse relationship between the housing price index and unemployment.

The correlation between poverty, the proportion of the population below the poverty line, and unemployment in Washington is 0.5519993. The correlation between poverty and unemployment is that of a large positive correlation. A large positive correlation suggests a strong positive relationship between the two variables that tend to move in the direction of one another. 

The correlation between the population and unemployment in Washington is -0.1159072. The correlation between the population and unemployment is that of a small negative correlation. A small negative correlation suggests a weak inverse relationship between the two variables.  

The correlation between the S&P 500 and unemployment in Washington is -0.02637235. The correlation between the S&P 500 and unemployment is that of a small negative correlation. A small negative correlation suggests a weak inverse relationship between the two variables. 

The correlation between the log of the population and unemployment in Washington is -0.1258998. The correlation between the log of the population and unemployment is that of a small negative correlation. The small negative correlation suggests a weak inverse relationship between the two variables. 

The correlation between the log of the median household income and unemployment in Washington is -0.375168. THe correlation between the log of the median household income and unemployment is that of a medium negative Correlation. A medium negative correlation suggests a somewhat strong inverse relationship between the two variables. 

```{r WA boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(WA_data$value, main="WA Unemployment", sub=paste("Outlier rows: ", boxplot.stats(WA_data$value)$out))  # box plot for 'Unemployment'
```

```{r WA boxplot poverty}
boxplot(WA_data$poverty, main="WA Poverty", sub=paste("Outlier rows: ", boxplot.stats(WA_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```

```{r WA linear regression all}
#run multiple linear model for data
WA_reg1 <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = WA_data)
summary(WA_reg1)
anova(WA_reg1)
```

```{r WA linear regression log all}
WA_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI, data = WA_data)
summary(WA_reg2)
anova(WA_reg2)
```

```{r WA linear significant variables}
WA_linearModelSignificant <- lm(value ~ poverty + log_pop, data = WA_data)
summary(WA_linearModelSignificant)
```

```{r OR Analysis}
OR_data <- merged_final[ merged_final$state == "OR", ]
```

```{r OR Descriptive Statistics}
#Descriptive statistics
stat.desc(OR_data)
```

The dependent variable, unemployment, has a mean of 6.46 and a standard deviation of 1.90. The lowest unemployment rate was in 2017 at 3.9, while the highest was in 2009 at 10.8. The mean HPI was 287.78, with a standard deviation of 137.46. The lowest HPI was in 1985 at 97.18, while the highest was in 2018 at 566.63. Poverty rate had a mean of 12.25 and a standard deviation of 1.40. In 1990, lowest rate was seen of 9.2, and the highest was 15.0 in 1998. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 15.04, and the standard deviation was 0.14. The lowest logged population at 14.796 in 1984, and the highest in 2018 at 15.25. Logged real median household income had a mean of 10.96 and a standard deviation of 0.071. In 1985, Oregon had the lowest log_RMHI at 10.80, while had the highest in 2018 at 11.14. 

```{r OR Pairwise Correlations}
#pairwise correlations
sapply(OR_data, class)
sapply(OR_data, is.factor)
cor(OR_data[sapply(OR_data, function(x) !is.character(x))])
```


```{r OR Histogram Unemployment}
hist(OR_data$value,
     main="Histogram of Unemployment from 1984-2018",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Oregon, from 1984 to 2018, the annual unemployment rates ranged from 2.0%-11.9%.The histogram did not produce a distribution with a skew worth acknowledging, however, the two most frequent ranges were 4.0-4.9% & 5.0-5.9% consecutively. There was only one year with an unemployment rate of 10.0% or higher.

```{r OR Histogram Poverty}
hist(OR_data$poverty,
     main="Histogram of Poverty from 1984-2018",
     xlab="Poverty Rate",
     breaks = 5)
```
In Oregon, from 1984 to 2018, the annual unemployment rates ranged from 9.0%-14.9%. The most frequent range of unemployment rates was 11.0-11.9% which occurred 15 of the 36 years. 

```{r OR scatter unemployment}
scatter.smooth(OR_data$year, OR_data$value, main="OR Unemployment by Year", xlab = "Year", ylab = "Unemployment", span = 2/8)
```

```{r OR scatter HPI}
scatter.smooth(OR_data$year, OR_data$HPI, main="OR HPI by Year", xlab = "Year", ylab = "HPI", span = 2/8)
```

```{r OR scatter poverty}
scatter.smooth(x=OR_data$year, y=OR_data$poverty, main="OR % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r OR scatter RMHI}
OR_data$RMHI <- as.integer(OR_data$RMHI)
scatter.smooth(x=OR_data$year, y=OR_data$RMHI, main="OR Median Household Income by Year", xlab = "Year", ylab = "RMHI", span = 2/8 )
```

```{r OR scatter population}
OR_data$population <- as.integer(OR_data$population)
scatter.smooth(x=OR_data$year, y=OR_data$population, main="OR Population by Year", xlab = "Year", ylab = "Population", span = 2/8 )
```

```{r OR scatter log RMHI}
#OR_data$log_RMHI <- as.integer(OR_data$log_RMHI)
scatter.smooth(x=OR_data$year, y=OR_data$log_RMHI, main="OR Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r OR scatter log population}
#OR_data$log_pop <- as.integer(OR_data$log_pop)
scatter.smooth(x=OR_data$year, y=OR_data$log_pop, main="OR Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r OR correlations}
cor(OR_data$RMHI, OR_data$value)
cor(OR_data$HPI, OR_data$value)
cor(OR_data$poverty, OR_data$value)
cor(OR_data$population, OR_data$value)
cor(OR_data$sp500, OR_data$value)
cor(OR_data$log_pop, OR_data$value)
cor(OR_data$log_RMHI, OR_data$value)
#run base correlations between categories and dependent variable
```

The correlation between the median household income and unemployment in Oregon is -0.3950978. The correlation between the median household income and unemployment is that of an inverse relationship. The inverse relationship is not the strongest, but does shows that in theory that as the median household income goes up, then the unemployment would go down. Of course, When it comes to scenarios such as a pandemic or depression, this could cause the relationship to become much more skewed, then it would be in most cases. 

The correlation between the housing price index and unemployment in Oregon is -0.07066041. The correlation between the housing price index and unemployment is an extremely weak relationship. With the correlation being that close to 0, it shows that there is a possibility of their being no correlation between the two variables. 

The correlation between poverty, the percentage of the population below the poverty line, and unemployment in Oregon is 0.4477392.  The correlation between poverty and unemployment is a decently strong positive relationship. This correlation suggests that the two variables tend to head in the same direction. As one goes up the other tends to do the same. 

The correlation between the population and unemployment in Oregon is 0.03283668. The correlation between population and unemployment is a weak positive correlation. With the correlation being so close to 0, the correlation suggests that there is no relationship between the two variables.

The correlation between the S&P 500 and unemployment in Oregon is -0.09391935. The correlation between the S&P 500 and unemployment is a weak negative relationship. A negative correlation suggests an inverse relationship between the two variables. But, the inverse relationship is extremely weak due to the closeness of the correlation being to 0. 

The correlation between the log of the population and unemployment in Oregon is 0.03392033. The correlation between the log of the population and unemployment is a weak positive correlation. With the correlation being so close to 0, it shows that relationship between the two variables is weak and does not exactly distinguish the direction that the two variables can go. 

The correlation between the log of the median household income and unemployment in Oregon is -0.3887452. The correlation between the log of the median household income and unemployment is a seemingly significant negative relationship. The negative relationship of correlation suggests that the two variables tend to move in opposite directions of each other. 

```{r OR boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(OR_data$value, main="OR Unemployment", sub=paste("Outlier rows: ", boxplot.stats(OR_data$value)$out))  # box plot for 'Unemployment'
```

```{r OR boxplot poverty}
boxplot(OR_data$poverty, main="OR Poverty", sub=paste("Outlier rows: ", boxplot.stats(OR_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```

```{r OR linear regression all}
#run multiple linear model for data
OR_reg1 <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = OR_data)
summary(OR_reg1)
anova(OR_reg1)
```

```{r OR linear regression log all}
OR_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI, data = OR_data)
summary(OR_reg2)
anova(OR_reg2)
```

```{r OR linear significant variables}
OR_linearModelSignificant <- lm(value ~ poverty + log_pop, data = OR_data)
summary(OR_linearModelSignificant)
```

```{r CO Analysis}
CO_data <- merged_final[ merged_final$state == "CO", ]
```

```{r CO Descriptive Statistics}
#Descriptive statistics
stat.desc(CO_data)
```

The dependent variable, unemployment, has a mean of 5.11 and a standard deviation of 1.82. The lowest unemployment rate was in 2000 at 2.5, while the highest was in 2010 at 8.8. The mean HPI was 279.72, with a standard deviation of 123.56. The lowest HPI was in 1988 at 123.59, while the highest was in 2018 at 570.27. Poverty rate had a mean of 10.56 and a standard deviation of 1.58. In 1997, lowest rate was seen of 8.2, and the highest was 13.7 in 1990. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 15.30, and the standard deviation was 0.19. The lowest logged population at 14.97 in 1984, and the highest in 2018 at 15.56. Logged real median household income had a mean of 11.09 and a standard deviation of 0.099. In 1984, Missouri had the lowest log_RMHI at 10.87, while had the highest in 2000 at 11.25. 

```{r CO Pairwise Correlations}
#pairwise correlations
sapply(CO_data, class)
sapply(CO_data, is.factor)
cor(CO_data[sapply(CO_data, function(x) !is.character(x))])
```


```{r CO Histogram Unemployment}
hist(CO_data$value,
     main="Histogram of Unemployment from 1984-2018",
     xlab="Unemployment Rate",
     breaks = 5)
```
In Colorado, from 1984 to 2018, the annual unemployment rates ranged from 2.0% to 8.9%. 11 of the 36 years were characterized by unemployment rates between the range of 5.0-5.9%. 

```{r CO Histogram Poverty}
hist(CO_data$poverty,
     main="Histogram of Poverty Across from 1984-2018",
     xlab="Poverty Rate",
     breaks = 5)
```
In Colorado, from 1984 to 2018, the poverty rates ranged from 8.0% to 13.9%. The histogram displays a slightly positively skewed distribution. 24 of the 36 years were characterized by poverty rates in the three ranges with the lowest values, 8.0-8.9%, 9.0-9.9%, and 10.0-10.9%.

```{r CO scatter unemployment}
scatter.smooth(CO_data$year, CO_data$value, main="CO Unemployment by Year", xlab = "Year", ylab = "Unemployment", span = 2/8)
```

```{r CO scatter HPI}
scatter.smooth(CO_data$year, CO_data$HPI, main="CO HPI by Year", xlab = "Year", ylab = "HPI", span = 2/8)
```

```{r CO scatter poverty}
scatter.smooth(x=CO_data$year, y=CO_data$poverty, main="CO % of People Below Poverty Level", xlab = "Year", ylab = "Poverty", ylim = c(0,20), span = 2/8)
```

```{r CO scatter RMHI}
CO_data$RMHI <- as.integer(CO_data$RMHI)
scatter.smooth(x=CO_data$year, y=CO_data$RMHI, main="CO Median Household Income by Year", xlab = "Year", ylab = "RMHI", span = 2/8 )
```

```{r CO scatter population}
CO_data$population <- as.integer(CO_data$population)
scatter.smooth(x=CO_data$year, y=CO_data$population, main="CO Population by Year", xlab = "Year", ylab = "Population", span = 2/8 )
```

```{r CO scatter log RMHI}
#CO_data$log_RMHI <- as.integer(CO_data$log_RMHI)
scatter.smooth(x=CO_data$year, y=CO_data$log_RMHI, main="CO Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r CO scatter log population}
#CO_data$log_pop <- as.integer(CO_data$log_pop)
scatter.smooth(x=CO_data$year, y=CO_data$log_pop, main="CO Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r CO correlations}
cor(CO_data$RMHI, CO_data$value)
cor(CO_data$HPI, CO_data$value)
cor(CO_data$poverty, CO_data$value)
cor(CO_data$population, CO_data$value)
cor(CO_data$sp500, CO_data$value)
cor(CO_data$log_pop, CO_data$value)
cor(CO_data$log_RMHI, CO_data$value)
#run base correlations between categories and dependent variable
```
The correlation between the median household income and unemployment in Colorado is -0.3744386. The correlation between the median household income and unemployment is a medium negative correlation. The correlation suggests an inverse relationship between the two variables. 

The correlation between the housing price index and unemployment in Colorado is -0.216168. The correlation between the housing price index and unemployment shows a weak negative correlation between the variables. The negative correlation illustrates the possibility of there being an inverse relationship between the two variables. The closeness of the correlation to 0 does show that there is a possibility that maybe the variables do not have any correlation at all. Of course with all of the depressions that have occurred and the pandemic this year, this could skew the correlation and downplay the relationship between the two variables. 

The correlation between poverty, the proportion of the population below the poverty line, and unemployment in Colorado is 0.7235819. The correlation between poverty and unemployment is a large positive correlation. A strong positive correlation suggests that there is a relationship between the two variables and it suggests that the two variables tend to move in the same direction of one another. 

The correlation between the population and unemployment in Colorado is -0.09548704. The correlation between the population and unemployment is a small negative correlation. The relationship between the two variables is quite small due to the correlation being so close to 0. 

THe correlation between the S&P 500 and unemployment is 0.0164226. The correlation between the S%P 500 and unemployment is an extremely small correlation. With the correlation being so close to 0, it signifies that the relationship between the two variables is not that common. Likewise, the correlation shows that there is a chance that there is no direct relationship between the two variables. 

The correlation between the log of the population and unemployment in Colorado is -0.1058047. The correlation between the log of the population and unemployment is a weak negative correlation. Just like in the case of the S&P 500, the weak correlation does not show a strong inverse relationship between the two variables. 

The correlation between the log of the median household income and unemployment in Colorado is -0.3621963. THe correlation between the log of the median household income and unemployment is that of a medium negative correlation. The correlation between these two variables in Colorado is much stronger than the S&P 500 and the log of the median household income. With the negative relationship, it suggests that the two variables tend to move in an inverse fashion. 
```{r CO boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(CO_data$value, main="CO Unemployment", sub=paste("Outlier rows: ", boxplot.stats(CO_data$value)$out))  # box plot for 'Unemployment'
```

```{r CO boxplot poverty}
boxplot(CO_data$poverty, main="CO Poverty", sub=paste("Outlier rows: ", boxplot.stats(CO_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```

```{r CO linear regression all}
#run multiple linear model for data
CO_reg1 <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = CO_data)
summary(CO_reg1)
anova(CO_reg1)
```

```{r CO linear regression log all}
CO_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI, data = CO_data)
summary(CO_reg2)
anova(CO_reg2)
```

```{r CO linear significant variables}
CO_linearModelSignificant <- lm(value ~ poverty + log_pop, data = CO_data)
summary(CO_linearModelSignificant)
```

>>>>>>> 32a2b1c8121c34957d746e37902519e2611353fb
```{r WV Analysis}
WV_data <- merged_final[ merged_final$state == "WV", ]
```

```{r WV Descriptive Statistics}
#Descriptive statistics
stat.desc(WV_data)
```

The dependent variable, unemployment, has a mean of 7.25 and a standard deviation of 2.48. The lowest unemployment rate was in 2006 at 4.2 while the highest was in 1984 at 14.40. The mean HPI was 164.11, with a standard deviation of 48.67. The lowest HPI was in 1984 at 87.28, while the highest was in 2018 at 230.86. Poverty rate had a mean of 17.60 and a standard deviation of 2.42. In 2004, lowest rate was seen of 14.2, and the highest was 22.4 in 1986. The sp500 index mean was 9.48, and the standard deviation was 15.70. In 2008, the lowest index was seen at -38.4, while the highest index was seen in 1995 at 34.11. The logged population mean was 14.42, and the standard deviation was 0.016. The lowest logged population at 14.40 in 1990, and the highest in 1984 at 14.47. Logged real median household income had a mean of 10.67 and a standard deviation of 0.10. In 1992, West Virginia had the lowest log_RMHI at 10.48, while had the highest in 2007 at 10.84.

```{r WV Pairwise Correlations}
#pairwise correlations
sapply(WV_data, class)
sapply(WV_data, is.factor)
cor(WV_data[sapply(WV_data, function(x) !is.character(x))])
```


```{r WV Histogram Unemployment}
hist(WV_data$value,
     main="Histogram of Unemployment from 1984-2018",
     xlab="Unemployment Rate",
     breaks = 5)
```

In West Virginia,from 1984 to 2018, the unemployment rates ranged from 4.0-15.9%. The histogram displays a positively skewed distribution. 15 of the 36 years were characterized by unemployment rates ranging in lowest interval, 4.0-5.9%. 

```{r WV Histogram Poverty}
hist(WV_data$poverty,
     main="Histogram of Poverty from 1984-2018",
     xlab="Poverty Rate",
     breaks = 5)
```

In West Virginia, from 1984 to 2018, the annual poverty rates ranged from 14.0% to 23.9%. This is the highest range out of all 8 states examined. Although this range has larger values comparatively, the histogram displays a positively skewed distribution. 25 out of the 36 years were characterized by unemployment rates that fell into the first two intervals, 14.0-15.9% & 16-17.9%.  

```{r WV scatter unemployment}
scatter.smooth(WV_data$year, WV_data$value, main="WV Unemployment by Year", xlab = "Year", ylab = "Unemployment", span = 2/8)
```

Unemployment in the state of West Virginia from 1984-2007 was at a declining pace, which shows that the job market is saturated and the state is in good standing. But, prior to the stock market crash of 2008-2009, we can notice that the unemployment rate began to start creeping up and shot up drastically when the stock market crash occurred. Since, then we have seen a steady decline to the rates that were prior to that of the stock market crash. 

```{r WV scatter HPI}
scatter.smooth(WV_data$year, WV_data$HPI, main="WV HPI by Year", xlab = "Year", ylab = "HPI", span = 2/8)
```

The housing price index in the state of West Virginia is much different than that of the poverty level and household income. THe housing price index has been steadily increasing within the state since 1984. THe one dip in the housing price index was the stock market crash of 2008-2009. As the scatter plot shows the housing price index dropped down for about two years and since then has continued to increase at a steady pace.

```{r WV scatter poverty}
scatter.smooth(x=WV_data$year, y=WV_data$poverty, main="WV % of People Below Poverty Level", xlab = "Year", ylab = "Poverty",  span = 2/8)
```

The poverty, proportion of the population below the poverty line, in West Virginia is pretty consistent with the other scatter plots related to the other variables. As we can in the scatter plot the recession of the 1990s caused the unemployment to shoot up as did the stock market crash in 2008-2009. The stock market crash affected the poverty level negatively for the next few years until the stock market was able to rebound and that is when the proportion of the population below the poverty line began to dwindle. 


```{r WV scatter log RMHI}
#WV_data$log_RMHI <- as.integer(WV_data$log_RMHI)
scatter.smooth(x=WV_data$year, y=WV_data$log_RMHI, main="WV Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

As the scatter plot suggests for the state of West Virginia, the median household income does fluctuate from year to year, but also shows a consistent increasing trend throughout the years. The dips on the median household income seem to come at time of economic recessions. In the 1990s, the recession caused a drop in the median household income. Additionally, the stock market crash that occurred in 2008-2009 caused a drop in the median household income. Other than those two instances the median household income within the state has increased steadily. 

```{r WV scatter log population}
#WV_data$log_pop <- as.integer(WV_data$log_pop)
scatter.smooth(x=WV_data$year, y=WV_data$log_pop, main="WV Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

The log of the population scatter plot for the state of West Virginia shows a steady decrease from 1984-1992. There started to be a slight increase in the population until about 1995 where it began to dwindle down again until 2001. From 2001, the population continued to increase steadily until 2012 and has been declining slightly ever since then.

```{r WV correlations}
cor(WV_data$RMHI, WV_data$value)
cor(WV_data$HPI, WV_data$value)
cor(WV_data$poverty, WV_data$value)
cor(WV_data$population, WV_data$value)
cor(WV_data$sp500, WV_data$value)
cor(WV_data$log_pop, WV_data$value)
cor(WV_data$log_RMHI, WV_data$value)
#run base correlations between categories and dependent variable
```

The correlation between the median household income and unemployment in West Virginia is -0.6170457. The correlation between the median household income and unemployment is that of a large negative correlation. A large negative correlation suggests that in most cases the variables tend to move in inverse directions. 

The correlation between the housing price index and unemployment in West Virginia is -0.7138695. The correlation between the housing price index and unemployment is that of a large negative correlation. Just like with the median household income, the housing price index suggests an inverse relationship between unemployment. 

The correlation between poverty, the proportion of the population below the poverty line, and unemployment in West Virginia is 0.6996082. The correlation between poverty and unemployment is that of a large positive correlation. A large positive correlation suggests that the two variables tend to move in the same direction unlike the median household income and the housing price index.

The correlation between the population and unemployment is 0.5144306. The correlation between the population and unemployment is that of a large positive correlation. Just like with poverty, the correlation suggests that the two variables tend to move in the same direction of one another. 

The correlation between the S&P 500 and unemployment in West Virginia is 0.1979507. The correlation between the S&P 500 and unemployment is a small positive correlation. The correlation shows that there is a small positive relationship between the two variables signifying that the two variables tend to move in the same direction.  

The correlation between the log of the population and unemployment in West Virginia is 0.509377. The correlation between the log of the population and unemployment is that of a large positive Correlation. The positive correlation suggests that the two variables mimic one another. 

The correlation between the log of the median household income and unemployment in West Virginia is -0.6322822. THe correlation between the log of the median household income and unemployment is a large negative correlation. The large negative correlation suggests that as the log of the median household income increases, the unemployment tends to decrease. 

```{r WV boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(WV_data$value, main="WV Unemployment", sub=paste("Outlier rows: ", boxplot.stats(WV_data$value)$out))  # box plot for 'Unemployment'
```

The box plot for the unemployment in the state of West Virginia shows that the median or second quartile for the unemployment rate within the state to be about 7%. Between the first and third quartile range we find the first quartile to be 5% and the third quartile to be at around 9%. This is where most of the unemployment between the years 1984 and 2019 tend to recide. The outliers for unemployment occur at times of economic instability that are found throughout the entire United States.


There is one outlier in the data based upon the boxplot, 1985 with an unemployment rate of 14.4%.
```{r WV boxplot poverty}
boxplot(WV_data$poverty, main="WV Poverty", sub=paste("Outlier rows: ", boxplot.stats(WV_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```

The poverty level within the state of West Virginia for the second quartile range is around 17%. THe first quartile begins at about 15.5% and the third quartile range is at 19%. The outliers for the data are similar to the outliers of the unemployment rate which occur at times of economic instability within the state, but also the entire United States. 

```{r WV linear regression all}
#run multiple linear model for data
WV_reg1 <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = WV_data)
summary(WV_reg1)
anova(WV_reg1)
```

```{r WV linear regression log all}
WV_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI, data = WV_data)
summary(WV_reg2)
anova(WV_reg2)
```

```{r WV linear significant variables}
WV_linearModelSignificant <- lm(value ~ poverty + log_pop, data = WV_data)
summary(WV_linearModelSignificant)
```


<<<<<<< HEAD
=======
```{r ME scatter RMHI}
ME_data$RMHI <- as.integer(ME_data$RMHI)
scatter.smooth(x=ME_data$year, y=ME_data$RMHI, main="ME Median Household Income by Year", xlab = "Year", ylab = "RMHI", span = 2/8 )
```

```{r ME scatter population}
ME_data$population <- as.integer(ME_data$population)
scatter.smooth(x=ME_data$year, y=ME_data$population, main="ME Population by Year", xlab = "Year", ylab = "Population", span = 2/8 )
```

```{r ME scatter log RMHI}
#ME_data$log_RMHI <- as.integer(ME_data$log_RMHI)
scatter.smooth(x=ME_data$year, y=ME_data$log_RMHI, main="ME Log Median Household Income by Year", xlab = "Year", ylab = "log_RMHI", span = 2/8 )
```

```{r ME scatter log population}
#ME_data$log_pop <- as.integer(ME_data$log_pop)
scatter.smooth(x=ME_data$year, y=ME_data$log_pop, main="ME Log Population by Year", xlab = "Year", ylab = "log_Population", span = 2/8)
```

```{r ME correlations}
cor(ME_data$RMHI, ME_data$value)
cor(ME_data$HPI, ME_data$value)
cor(ME_data$poverty, ME_data$value)
cor(ME_data$population, ME_data$value)
cor(ME_data$sp500, ME_data$value)
cor(ME_data$log_pop, ME_data$value)
cor(ME_data$log_RMHI, ME_data$value)
#run base correlations between categories and dependent variable
```
The correlation between the median household income and unemployment is Maine is -0.1586851. THe correlation between the median household income and unemployment is a small negative relationship. The negative relationship suggests that the two variables tend to move inversely of one another. 

The correlation between the housing price index and unemployment in Maine is -0.02225343. The correlation between the two variables is that of a small negative relationship. The relation between the two variables is a small correlation that points to the two variables moving inversely of one another. 

The correlation between poverty, the proportion of the population below the poverty line, and unemployment in Maine is 0.3808906. The correlation between poverty and unemployment is that of a medium positive correlation. The positive correlation suggests that the two variables tend to mimic one another with the direction that they proceed.

The correlation between the population and unemployment in Maine is 0.06660653. The correlation between the population and unemployment is that of a small positive correlation. A small positive correlation suggests that there is a weak positive relationship between the population and unemployment, but the two variables do tend to move in the same direction of one another. 

The correlation between the S&P 500 and unemployment in Maine is 0.03233341. The correlation between the S&P 500 and unemployment is a small positive correlation. The small positive correlation is quite close to 0, but does still show the small possibility that the two variables move in the same direction of one another. 

The correlation between the log of the population and unemployment in Maine is 0.06676556. The correlation between the log of the population and unemployment is quite similar to that of the S&P 500. The correlation is a small positive correlation between the two variables. 

The correlation between the log of the median household income and unemployment in Maine is -0.1587006. The correlation between the log of the median household income and unemployment is that of a small negative relationship. The small negative relationship suggests that the two variables have a weak inverse relationship of one another. 

```{r ME boxplot unemployment}
par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(ME_data$value, main="ME Unemployment", sub=paste("Outlier rows: ", boxplot.stats(ME_data$value)$out))  # box plot for 'Unemployment'
```

```{r ME boxplot poverty}
boxplot(ME_data$poverty, main="ME Poverty", sub=paste("Outlier rows: ", boxplot.stats(ME_data$poverty)$out))  # box plot for 'Poverty'
#create box plot for poverty level and unemployment level
```

```{r ME linear regression all}
#run multiple linear model for data
ME_reg1 <- lm(value ~ poverty + RMHI + HPI + population + sp500, data = ME_data)
summary(ME_reg1)
anova(ME_reg1)
```

```{r ME linear regression log all}
ME_reg2 <- lm(value ~ HPI + poverty + sp500 + log_pop + log_RMHI, data = ME_data)
summary(ME_reg2)
anova(ME_reg2)
```

```{r ME linear significant variables}
ME_linearModelSignificant <- lm(value ~ poverty + log_pop, data = ME_data)
summary(ME_linearModelSignificant)
```
>>>>>>> 32a2b1c8121c34957d746e37902519e2611353fb

Next, the data we collected is considered a time series as it was measured on an annual basis.  We used the unemployment data that was collected for each state and implemented a time series forecast using the Holt Method.  

Time series data typically contains four components: trend, seasonal, cyclical, and random components.  Trend is represented by the long-term movements in a data series, whether that be upward or downward.  Seasonality is normally represented by repetitions that occur within a single year time period, whereas the cyclical component represents long-term trends that are usually a factor the economy.  Typically, seasonal trends are easier to identify than cyclical trends because the time period of a seasonal trend is often known ahead of time.  Random components represent the random and unexplained movements within the time series data.

As the unemployment data we collected is shown on an annual basis, we will be focusing on the trend, cyclical, random components in our analysis.  The type of analysis we have chosen to focus on is called Holt Exponential Smoothing Method.  This method is characterized by incorporating large upward and downward fluctuations in data series.  This method is especially relevant because the dataset lacks seasonal variation.  

```{r IL Holt Analysis}

IL_TS <- ts(IL_data$value, start = c(1984), end = c(2018), frequency = 1)
#Stores start, end, and frequency of timeseries data

IL_TData <- window(IL_TS, end = c(2004))
IL_VData <- window(IL_TS, start = c(2005))
#Partitions the training and validation sets
```

In the above section of code, we begin our Holt analysis of IL by creating the time series with the TS function.  We set the start and end values of our times series data equal to the time span our data stretches and set the frequency equal to one.

Using the window function we partition our dataset into a training set, IL_TData, and a validation set, IL_VData.

```{r IL Holt Analysis User}
IL_HUser <- ets(IL_TData, model = "AAN", alpha = 0.2, beta = 0.15)
summary(IL_HUser)
#Training data set using user based parameters for alpha and beta for smoothing.
```

It is common for practitioners to provide user supplied values for alpha and beta (smoothing parameters) due to the fact that the computer generated values are known to over-fit models where the data performed well in the sample period, but does not exhibit the same performance in the future.  In our data, we compare both user supplied and computer generated smoothing parameters.  We then compare the error measures to determine which model is a better fit to our data.  We used an alpha = 0.2 and beta = 0.15 when creating the user supplier parameters model.

The ets function above denotes the error, trend, and seasonality of the training set. The first letter in the model string represents that we want the error to be additive, the second letter represents that fact that we expect the trend type to be additive, and the third letter, "N", represents the fact that we do not intend on incorporating seasonality into our model.


```{r IL Holt Analysis Computer}
IL_HCmp <- ets(IL_TData, model = "AAN")
summary(IL_HCmp)
#Training data set using computer based parameters for alpha and beta for smoothing.
```

We replicated the ets function using the computer generated smoothing parameters and obtained the output above. By comparing the root mean squared error(RMSE) and the mean absolute percentage error(MAPE) we can assess the fitness of the user supplier and computer generate models.  The user supplied model has a higher RMSE and MAPE indicating that the computer generated model provides a better overall fit for making a forecast. 

```{r IL Holt Set Forecast}
IL_nV <- length(IL_VData)
IL_fUser <- forecast(IL_HUser, h = IL_nV)
IL_fCmp <- forecast(IL_HCmp, h = IL_nV)
```

The length function is used to set the number of observations equal to the length of the validation set.
Then the forecast function is utilized to make a forecast based on the observations in our training set.

```{r IL Holt Analysis User Accuracy}
accuracy(IL_fUser,IL_VData)
```
```{r IL Holt Analysis Computer Accuracy}
accuracy(IL_fCmp,IL_VData)
```

The accuracy function is used to test the error term outputs that were created using the training set.  When comparing the accuracy function results we see conflicting results in our validation set.  The test set shows that the RMSE and MAPE of the user generated model provides a better fit for the performance of our data.  Ultimately, we decided to continue the process using the computer generated model, but the user supplied model would work just as well.
  
``` {r IL Holt Forecast}
IL_HFinal <- ets(IL_TS, model = "AAN")
forecast(IL_HFinal, h=1)
```

Finally, everything is brought together with the ets and forecast functions to forecast the unemployment level for the next period, 2019. The forecast generated shows that the expected value of the Illinois unemployment level in December 2019 is equal to 4.42% based on the trends that is observed in our model.  Based on the forecast we expect that the unemployment will range from 2.85% to 5.98% at a 80% confidence interval.  Additionally, we can be 95% confident that the unemployment value will fall between 2.02% and 6.81%. Based on the actual unemployment rate in IL for December 2019 of 3.5% our model was accurate at predicting the unemployment value at both the 95% and 80% confidence intervals.


```{r WV Holt Analysis}

WV_TS <- ts(WV_data$value, start = c(1984), end = c(2018), frequency = 1)
#Stores start, end, and frequency of timeseries data

WV_TData <- window(WV_TS, end = c(2004))
WV_VData <- window(WV_TS, start = c(2005))
#Partitions the training and validation sets
```
```{r WV Holt Analysis User}
WV_HUser <- ets(WV_TData, model = "AAN", alpha = 0.2, beta = 0.15)
summary(WV_HUser)
#Training data set using user based parameters for alpha and beta for smoothing.
```
```{r WV Holt Analysis Computer}
WV_HCmp <- ets(WV_TData, model = "AAN")
summary(WV_HCmp)
#Training data set using computer based parameters for alpha and beta for smoothing.
```

With West Virginia, we setup the datasets and partitioned them into training and validation sets following similar steps to Illinois.

The user supplied model for the state of West Virginia has a higher RMSE and MAPE at 1.4288 and 14.1598, respectively. The computer generated model has a RMSE of 0.9736 and a MAPE of 8.5906 indicating that the computer generated model provides a better overall fit to our forecast. 

```{r WV Holt Set Forecast}
WV_nV <- length(WV_VData)
WV_fUser <- forecast(WV_HUser, h = WV_nV)
WV_fCmp <- forecast(WV_HCmp, h = WV_nV)
```
```{r WV Holt Analysis User Accuracy}
accuracy(WV_fUser,WV_VData)
```
```{r WV Holt Analysis Computer Accuracy}
accuracy(WV_fCmp,WV_VData)
```

The accuracy function confirms that the computer generated model is a better overall fit to the data we are using to create the model.  This is due to the fact that once again the validation set exhibits a lower RMSE and MAPE under the computer generated model.

``` {r WV Holt Forecast}
WV_HFinal <- ets(WV_TS, model = "AAN")
forecast(WV_HFinal, h=1)
```

The forecast for West Virginia shows us that the expected value of the unemployment level in December 2019 is equal to 4.70% based on the trends that were observed in the computer generated model. At an 80% confidence interval we expect West Virginia's unemployment level to range from 3.20% to 6.20%.  Additionally, we can be 95% confident that the unemployment rate will be in the 2.41% to 6.99% level. The actual unemployment rate for the State of West Virginia in December 2019 is 5%. Our model is accurate at both the 95% and 80% confidence intervals.  5% is also fairly close to the point estimate at 4.7% further signifying that our model produces accurate results.